### 2014-04-25 article.tex
Change from Geophys. research letter tpl to custom one (based on geogg101)

\documentclass{article}

% Some useful packages.
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{mhchem}

% Reduces margins substantially.
\usepackage{geometry}
\newgeometry{margin=2.5cm}

% Allows headers and footers.
\usepackage{fancyhdr}
\pagestyle{fancy}
% Get rid of annoying line under header.
\renewcommand{\headrulewidth}{0pt}

\lhead{}
\chead{BTTM8 - CLIMATE MODELLING - USING HADAM3 TO MODEL CLIMATE CHANGE}
\rhead{}
%\lfoot{}
%\cfoot{}
%\rfoot{}

\usepackage[backend=bibtex,style=authoryear,sorting=nyt,dashed=false]{biblatex}
\bibliography{references/references}
\renewcommand*{\nameyeardelim}{\addcomma\space}

\begin{document}

\part*{Using HadAM3 to model the effects of climate change}

\section*{Abstract}

\ce{CO2} climate sensitivity (\SI{}{W.m^{-2}.K^{-1}})

\section{Introduction}

According to \textcite{gregory2004new}, you can never \dots

\section{Methods}

\subsection{Model}

\section{Results}

\begin{figure}[hbp]
    \centering
    \includegraphics[width=\textwidth]{figures/figure_1}
    \caption{Caption.}
    \label{fig:figure_1}
\end{figure}

\section{Discussion}

\section{Conclusions}

\appendix

\section{Required output}

\begin{enumerate}
    \item first
    \item first
    \item first
    \item second
    \item second
\end{enumerate}

\end{document}